---
layout: post
title: Clustering
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It’s quite common the case that one needs to cluster categorical variables.
When it comes to clustering, a number of well known techniques abound
such as *k-means*, *k-medoids*, and *hierarchical clustering*. With *k-means*
and *k-medoids* clustering, you must have an idea of the number of
clusters $k$ desired. Furthermore, the final clusters can vary
significantly depending on the number of initial clusters chosen. With
hierarchical clustering, an intuitive cluster hierarchy can be observed,
however computational efficiency and immediate utility is less clear.
Nonetheless, the hierarhical clustering approach lends itself well to
interpretation so let us consider this approach first. 

```{r}
d=dist(as.matrix(mtcars))
hc=hclust(d)
plot(hc)
```

We can further decide on an appropriate number of clusters, say 3, and observe membership.

```{r}
clustercut = cutree(hc,3)
sort(clustercut)
```

It is not always the case that cluster memberships break up so easily into evenly sized, appropriate clusters.  

For this case, we will use the [Markov Clustering Algorithm](http://micans.org/mcl/)  because this will allow us to distinguish between overlapping categories. 

Stijn van Dongen, Graph Clustering by Flow Simulation, PhD thesis, University of Utrecht, May 2000.
( http://www.library.uu.nl/digiarchief/dip/diss/1895620/inhoud.htm )

For example, consider a group of movies.  Movies are usually grouped under multiple genres such as drama, action, documentary, etc.  This makes it tricky to truly distinguish one movie from the next.  Perhaps there are more reprentative genre groupings.  A sample of our incoming dataset is given here.  This data comes courtesy of IMDb and is available [here](https://www.kaggle.com/deepmatrix/imdb-5000-movie-dataset).

```{r echo=FALSE}
df = read.csv('~/code/data/movie_metadata.csv',stringsAsFactors=FALSE)
```
```{r echo=TRUE}
head(df[,c('movie_title','genres')])
```

In order to use hierarchical clustering, we need to assign a distance value between every element and every other element. For $n$ elements, this usually involves needing to compute $(n^2-n)/2 = n(n-1)/2$ values.  In our list we have `r rows=dim(df)[1]; rows` movies.  So we need to do `r format((rows^2-rows)/2, scientific=TRUE)` computations.  To simplify this task we will sample 100 movies, a more reasonable level. 
```{r}
set.seed(123)
sample_size = 100
sample_index = sample(dim(df)[1], sample_size, replace=FALSE)
df_sample = df[sample_index,]
```
One simple measure is the Jaccard index: the intersection of sets over the union of sets.  
That is, given movies $A$ and $B$,
$$J(A,B) = \frac{|A\cap B|}{|A\cup B|}$$
Where the absolute values represent the size of the sets in question.  We can preallocate a vector to represent distance values.
```{r}
distances = rep(0,(sample_size^2-sample_size)/2)
```
Then we simply need to iterate over the movies in our sample set to see what the Jaccard index of any pair of movies is according to genre tags.
```{r}
npairs = (sample_size^2-sample_size)/2
distances = rep(0,npairs)
movie_titlesA = rep(" ", npairs)
movie_titlesB = rep(" ", npairs)
k=0
for (i in 1:(sample_size-1)){
  for (j in (i+1):sample_size){
    k=k+1
    A = unlist(strsplit(df_sample[i,'genres'],"[| ]"))
    B = unlist(strsplit(df_sample[j,'genres'],"[| ]"))
    movie_titlesA[k] = df_sample[i,'movie_title']
    movie_titlesB[k] = df_sample[j,'movie_title']
    distances[k]= length(intersect(A,B))/length(union(A,B))
  }
}
df_out = data.frame(cbind(movie_titlesA,movie_titlesB,distances),stringsAsFactors=FALSE)
df_out = df_out[df_out$distances > 0,]
```
After we export the dataframe to a file `
```{r echo=FALSE}
write.table(df_out,file="~/code/data/movie_pairs.abc",sep="\t",quote=FALSE,row.names=FALSE,col.names=FALSE)
```
We call the mcl algorithm `mcl movie_pairs.abc --abc -I 4.0`.  `4.0` is the extent to which we would like the algorithm to differentiate. This one parameter allows us to vary the final number of clusters we arrive at.

> The Nativity Story 	The Holy Girl 	Twin Falls Idaho 	Chloe 	The Age of Adaline 	Queen of the Mountains 	Illuminata 	The Clan of the Cave Bear 	Remember the Titans 	Forsaken 	The End of the Affair 	Everest 	Courage 	The Crow 	Dracula Untold 	A Farewell to Arms 	The Celebration 	The Passion of the Christ 	All or Nothing 	The Express 	Day One 	Adam 	Solomon and Sheba 	The Outsiders 	Freeheld 	Mooz-Lum 	Evita 

> Cluster 1: Elysium 	The Scorpion King 	X-Men Origins: Wolverine 	The Mask 	London Has Fallen 	In the Valley of Elah 	N-Secure 	Airlift 	Grosse Pointe Blank 	Spider-Man 3 	The Heat 	Ride Along 	Beverly Hills Cop II 	The Living Daylights 	Wing Commander 	12 Rounds 	The Dead Undead 	Fast Five 	Desperado 	The Dark Knight Rises 

> Cluster 2: But I'm a Cheerleader 	Bran Nue Dae 	Tumbleweeds 	Shortbus 	Friday 	Agora 	Quigley Down Under 	Sonny with a Chance   Coyote Ugly 	Bandits 	The Intern 	The Love Letter 	Of Horses and Men 	Cold Mountain 	Topsy-Turvy 	Annie Get Your Gun 

> Cluster 3: Connie and Carla 	Harold & Kumar Go to White Castle 	For Your Consideration 	Airplane! 	Monster-in-Law 	Team America: World Police 	The Big Bounce 	Corky Romano 	Supporting Characters 	The Wash 	An Alan Smithee Film: Burn Hollywood Burn 	Chicago 

> Cluster 4: The Adventurer: The Curse of the Midas Box 	The Secret of Kells 	Monty Python and the Holy Grail 	Chicken Little 	Stuart Little 2 	Big Fat Liar 	Yogi Bear 	The Wild Thornberrys Movie 	Journey 2: The Mysterious Island 	The Last Legion 	Dragon Hunters 

> Cluster 5: The Call of Cthulhu 	The Mothman Prophecies 	A Nightmare on Elm Street 2: Freddy's Revenge 	Premonition 	The Jacket 	Urban Legend 	The Visit 	Red Eye 	Thir13en Ghosts 

> Cluster 6: The Adjustment Bureau 	The Invasion 	Beyond the Black Rainbow 

> Censored Voices 	My Date with Drew 
