---
layout: post
title: Canonicalization
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It’s quite common the case that I need to cluster categorical variables.
When it comes to clustering, a number of well known techniques abound
such as *k-means*, *k-medoids*, and *hierarchical clustering*. With *k-means*
and *k-medoids* clustering, you must have an idea of the number of
clusters $k$ desired. Furthermore, the final clusters can vary
significantly depending on the number of initial clusters chosen. With
hierarchical clustering, an intuitive cluster hierarchy can be observed,
however computational efficiency and immediate utility is less clear.
Nonetheless, the hierarhical clustering approach lends itself well to
interpretation so let us consider this approach first. For example,
consider a group of vehicles for transportation.

> Car, Motorboat, Hatchback, Sailboat, Scateboard, Motorcycle, Pickup Truck, Oil Tanker, Dump Truck, Bike, Electric Scooter

By no means is this an exhaustive list. Suppose we want to arrive at a
smaller grouping of variables encompassing the original list. In order
to use hierarchical clustering, we need to assign a distance value
between every element and every other element. For $n$ elements, this
usually involves needing to know $(n^2-n)/2 = n(n-1)/2$ elements.  In our
above list of $10$ elements, this would be composed of a $10x10$ upper/lower
triangular distance matrix where each element represent how close or far each element is to the next.
The matrix does not need to be fully populated because distances are assumed to be symmetric.  
Suppose we have such a matrix. We will let the rows and columns of the matrix be in the exact same order as above
in R.
```{r m}
m <- matrix(c(0,1,1,10,1,1,1,1,10,1,0,0,1,10,1,1,1,1,10,1,0,0,0,10,1,1,1,1,10,1,0,0,0,0,10,10,10,10,1,1,0,0,0,0,0,1,1,1,10,1,0,0,0,0,0,0,1,1,10,1,0,0,0,0,0,0,0,1,10,1,0,0,0,0,0,0,0,0,10,1,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0),nrow=10,ncol=10)
m
```

